### 2024-08-29 sports-cognitive-psychology.qmd
20240829 - copyediting

- Bill James, "Moneyball", Billy Beane, Theo Epstein, and role of Sabermetrics in helping MLB teams with limited budgets (e.g., Oakland Athletics) compete with larger budgets: using statistics to identify player value more accurately, especially for identifying undervalued players
- development of advanced metrics (e.g., value over replacement player, wins above replacement, fieling independent pitching)
- focus on on-base percentage and slugging percentage (and on-base plus slugging percentage) instead of batting average; defensive shifting; fewer attempts to steal bases; fewer bunts; batters taking more pitches; more frequent pitching changes (for pitcher–batter matchups, e.g., righty–lefty); focus on velocity and spin rate
- other sports, including basketball and football are making more use of analytics
- basketball: types of shots (focusing on three pointers and layups); star players resting more games
- football: going for it on fourth down; greater emphasis on the passing game; drafting Running Backs later in the draft (and, overall, valuing Running Backs less); trading down in the draft for more low picks than fewer high picks [because top picks are frequently overvalued; @Massey2013]

>>>>

- Bill James, "Moneyball", Billy Beane, Theo Epstein, and role of Sabermetrics in helping MLB teams with limited budgets (e.g., Oakland Athletics) compete with larger budgets: using statistics to identify player value more accurately, especially for identifying undervalued players
- development of advanced metrics (e.g., value over replacement player, wins above replacement, fielding independent pitching)
- focus on on-base percentage and slugging percentage (and on-base plus slugging percentage) instead of batting average; defensive shifting; fewer attempts to steal bases; fewer bunts; batters taking more pitches; more frequent pitching changes (for pitcher–batter matchups, e.g., righty–lefty); focus on velocity and spin rate
- other sports, including basketball and football are making more use of analytics
- basketball: types of shots (focusing on three pointers and layups); star players resting more games
- football: going for it on fourth down; greater emphasis on the passing game; drafting Running Backs later in the draft (and, overall, valuing Running Backs less); trading down in the draft for more low picks than fewer high picks [because top picks are frequently overvalued; @Massey2013]
